---
title: "Colleges"
author: "Harshil Kurapati"
format: html
execute:
  echo: false
---

```{r}
#| message: false
library(tidyverse)
library(primer.data)
library(tidymodels)
library(broom)
library(marginaleffects)
```

```{r}
#| cache: true
x <- colleges |>
  select(tuition, grad_rate, selectivity) |>
  filter(tuition > 2) |>
  drop_na() |>
  mutate(selectivity = factor(as.character(selectivity),
                              levels = c("Elite", "Highly Selective",
                                         "Moderately Selective",
                                         "Lowly Selective", "Non-selective")))
```

```{r}
#| cache: true
fit_colleges <- linear_reg() |>
  set_engine("lm") |>
  fit(grad_rate ~ tuition + selectivity, data = x)
```

```{r}
plot_predictions(fit_colleges, condition = c("tuition", "selectivity"), draw = FALSE) |>
  ggplot(aes(x = tuition, y = estimate, color = selectivity)) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = selectivity),
              alpha = 0.15, color = NA) +
  geom_line(linewidth = 1) +
  labs(
    title = "Predicted Graduation Rate by Tuition and Selectivity",
    subtitle = "Within each selectivity tier, every $10k of tuition adds ~9 percentage points",
    x = "Tuition ($10,000)",
    y = "Predicted Graduation Rate",
    color = "Selectivity",
    fill = "Selectivity",
    caption = "Source: IPEDS 2013 via primer.data"
  ) +
  theme_minimal()
```

Graduation rates vary widely across US colleges and are strongly associated with institutional tuition and selectivity. Using data from the Department of Education's 2013 IPEDS database we seek to estimate the difference in expected graduation rates between colleges with different tuition levels. One concern is that the 2013 data is over a decade old and the relationship between tuition and graduation rate may have shifted significantly since then as college pricing and financial aid models have changed. We model graduation rate as a linear function of tuition in units of $10,000 and selectivity tier with Elite colleges as the reference category. We estimate that within a given selectivity tier each $10,000 of tuition is associated with about 9 percentage points higher graduation rate with a 95% confidence interval of 7 to 11 percentage points.
